---
title: DATA-413/613 Homework 05 API Data
author: Your Name
number-sections: true
embed-resources: true
format: html
---



**Use the {tidyverse}, {keyring}, {httr2}, and {jsonlite} packages for this question.**

```{r}
#| message: false
library(tidyverse)
library(keyring)
library(httr2)
library(jsonlite)
```


Next steps 

Find sources on Why average american care about where lawmakers raise money

As we find them post links and a few sentence summary of what the article says 




#here's some website links we discussed  
#this is the api documentation
https://api.open.fec.gov/developers/#/
#this is a map that shows presidential donations 
https://www.fec.gov/data/candidates/president/presidential-map/
#this is an interactive method users can dig into to look at individual elections 
https://www.fec.gov/data/elections/?state=&cycle=2026&election_full=true
#this shows aggregate donations for candidates 
https://www.fec.gov/data/elections/senate/NC/2026/


#heres' an old project I did with my coworkers that looked at the Arizona senate race that might help us figure out how to join data to shapefiles 

https://github.com/ChrisCioffi/districts/blob/master/candidates.R

1.  Pick a website of your choice (not discussed in class) that requires a free API key and has recent data. You may use an R package for the website or use {httr2} to interact directly with the API.\

I chose https://api.open.fec.gov/developers/

```{r}
# key_set("fec_api_key")
```

2.  **Identify the link to instructions to obtain a key for the API.**\
3.  Use your API with {keyring} to download a data set with multiple variables. **Do not include your key in your file.**\
4.  Convert elements of interest into a tibble

```{r}
# Set the base URL for the FEC API
base_url <- "https://api.open.fec.gov/v1/candidates/totals/"
params <- list(election_year = "2026", state = "NC", per_page = 100, office = "S", election_full = TRUE, is_active_candidate = TRUE)

# Construct the API query (example: candidate data)
fundraising <- (request(base_url) |>
  req_url_query(!!!params, api_key = keyring::key_get("fec_api_key")) |>
  req_perform() |>
  resp_body_json(resp = _))$results |>
  tibble(data = _) |>
  unnest_wider(data)
```

5.  State a question of interest.

Is Wiley Nickel putting up a legit challenge to NC Sen. Thom Tillis?

6.  Create an appropriate plot with proper labels and theme to analyze your question of interest.

```{r}
options(scipen = 999)
fundraising |>
  filter(has_raised_funds == TRUE) |> # get rid of the candidates who haven't raised money
  rename(
    `Total receipts` = receipts,
    `Other political committee contribs.` = other_political_committee_contributions,
    `Contributions over $200` = individual_itemized_contributions
  ) |>
  select(name, `Total receipts`, `Other political committee contribs.`, `Contributions over $200`) |>
  pivot_longer(
    cols = -name, # Keep name column as is
    names_to = "contribution_type",
    values_to = "amount"
  ) |>
  ggplot(aes(x = name, y = amount, fill = contribution_type)) +
  geom_col(position = "dodge") + # Use position = "dodge" to make bars next to each other
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + # Rotate the x-axis labels for better readability
  labs(
    x = "Candidate",
    title = "Fundraising Contributions by Candidate",
    subtitle = "Tillis shows he has raised more and got more large-dolar contributions than Nickel",
  )
```



7.  Interpret the plot to answer your question.

Thom Tillis so far has a cash advantge over Wiley Nickel. Not only has he raised more, but he's getting more from other political committees and from donors who gave more than $200.

8.  Read the site guidance for use of their data, e.g., terms and conditions, API documentation, or FAQ (if they exist)) and, in a few sentences, comment on the severity or strictness of the guidance or the lack of guidance. Discuss how the guidance could enable or restrain unethical uses of the data.

One of the things that the FEC points out, is that users cannot use use contributor lists for commercial purposes or to solicit donations. That's particularly important, because if a donor has already given to candidates, they are probaly more likely to give to others.

What's interesting about how they police that, is the Federal Election Campaign Act (the Act) may use a method of detecting whether the names and addresses of individual contributors are being used illegally by “salting” the report with up to ten fictitious contributor names. "A portion of the committee's unitemized contributions is attributed to each of the fictitious contributors."

That could easily get someone in trouble if they solicit contributions to someone who the committee knows is fictitious, and they may may file a complaint with the FEC.


##### multiple pages 

took a generative AI solution for multiple pages - didn't go great 
this is what we will need to get running to get all the contributions and group them by state 

```{r}
library(httr2)
library(tidyverse)

# Set API parameters
base_url <- "https://api.open.fec.gov/v1/candidates/"
params <- list(
  election_year = "2026",
  office = "S",
  per_page = 100,
  election_full = TRUE,
  api_key = key_get("fec_api_key")
)

# Create the request
req <- request(base_url) |> 
  req_url_query(!!!params) |> 
  req_throttle(rate = 30 / 60)

# Use req_perform_iterative with built-in pagination
responses <- req_perform_iterative(req, 
                                   iterate_with_offset("page_index"))

# Extract JSON and combine into one tibble
fec_data <- responses |>
  map(resp_body_json, simplifyVector = TRUE) |> 
  map("results") |> 
  keep(~ length(.x) > 0) |> # Remove empty responses
  map_dfr(as_tibble) # Bind into one tibble

# View the dataset
glimpse(fec_data)

```

